---
title: "Final Project"
author: "Jenny Gu and Yijiang Zhao"
date: "May 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prevalence of Hate Crimes in Modern Society
We analyzed a data set which detailed all instances of hate crime reported to the **UCR** (Uniform Crime Reporting) and the **NIBRS** (National Incident-Based Reporting System) from 2000 until 2017. These crimes ranged anywhere from intimidation to violent crimes, and were based on various biases, such as race, gender, sexuality, religion, etc.

## Contingency Tables of Race of Perpetrator vs. Race of Victim
One important issue and concern was that surrounding race-based hate crime.
Using contingency tables, for both the actual and the expected frequencies, we found that the race of the perpetrator and the race of the victim are not independent.

<center>
``` {r}
# Observed
formattable(convert (freqs, tbl))
# Expected
formattable(convert (exps, Expected))
```
</center>

We performed a chi-squared test to further display this and found a p-value of <2.2E-16, which is statistically significant on all levels (i.e. 0.1, 0.05, and 0.01). Therefore, the race of the perpetrator and the race of the victim are not independent.

``` {r}
chisq.test(offender, victim) 
```

## How Much Hate?

One preliminary analysis we wanted to do was see how the distribution of the number of hate crimes in each city. To keep the overall population of a city constant, we only selected the hate crime data for cities with a population of 250,000 to 499,999 and created a histogram. We fitted a normal curve over the distribution and found that it was not normal. 

<center>
``` {r, echo=FALSE}
ggplot(citydf, aes(x=citydf$count)) + labs(title="Frequency of Hate Crimes in Cities 250,000 - 499,999", x="Number of Hate Crimes", y = "Density") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  geom_histogram(aes(y = ..density..), binwidth=1, fill="#244B70") + 
  stat_function(
    fun = dnorm, 
    args = with(citydf, c(mean = mean(citydf$count), sd = sd(citydf$count))),
    color = "red"
  )
```
</center>

We then overlaid a chi distribution (green), gamma distribution (red), and an exponential distribution (blue), finding that the gamma function fit best.

<center>
``` {r, echo=FALSE}
hist(HC[which(HC$desc == "Cities from 250,000 thru 499,999"),]$tot, breaks=130, freq=FALSE, xlab = "Number of Hate Crimes", main = "Frequency of Hate Crimes in Cities 250,000-499,999")

curve(dexp (x, rate=.15), add = TRUE, col = "blue", lwd = 2)
curve(dchisq(x, df=5), add=TRUE, col="green")
curve(dgamma(x, shape=5.6), add=TRUE, col="red")
```
</center>

However, by taking 10000 samples of 10 data points and averaging the total numbers of hate crime for each sample, the means of each of those samples better fit a normal distribution, per the central limit theorem. When plotting the normal distribution (red) and the gamma function (green), we found that the gamma function still fits slightly better.

<center>
```{r, echo=FALSE}
ggplot(avgHCdf, aes(x = avgHCdf$avgHC)) + 
  geom_histogram(aes(y = ..density..), binwidth=0.5, fill="#244B70") + 
  stat_function(
    fun = dnorm, 
    args = with(avgHCdf, c(mean = mean(avgHCdf$avgHC), sd = sd(avgHCdf$avgHC))),
    color = "red"
  ) + 
  stat_function(
    fun = dgamma, 
    args = with(avgHCdf, c(shape=10, rate = 0.83)),
    color = "green"
  ) + labs(title="Sampling Hate Crimes Frequencies for Cities 250,000-499,999", x="Number of Hate Crimes", y = "Density") + theme(plot.title = element_text(hjust = 0.5))
```
</center>

We find that it is slightly skewed right (with a factor of 0.66), though not as visibly so as actual distribution.

```{r}
skewness(avgHC, type=1)
```

## National Frequencies of Hate Crime Throughout Time
Another issue of concern was how hate crime changed over time. We first created a barplot which displayed the national counts of instances of hate crime for each year.

<center>
```{r histogram, echo=FALSE}

ggplot(data=data.frame(year, totCrimes), aes(x=year,y=totCrimes)) + geom_bar(stat="identity",fill="#244B70") + labs(title="Total Incidents of Hate Crime over Time", x="Year", y = "Total Hate Crimes Count") +
  theme(plot.title = element_text(hjust = 0.5))

```
</center>

We discovered that 2001 seemed somewhat higher in the number of hate crimes and performed statistical analysis and found that it was indeed significantly higher. <br><br>

``` {r}
# observed difference in 2001
obsDiff2001
N <- 10000
# mean of differences
mean2001
# pvalue
pvalue2001 <- (sum(abs(diffs) >= abs(obsDiff2001))+1)/(N+1)
pvalue2001
```

The p-value from the permutation test is statistically significant under all alpha levels (0.1, 0.05, 0.01).
<br><br>
To explore what might be driving this increase in hate crimes, we hypothesized that it could be the result of Anti-Islam and Anti-Arab hate crimes increasing as the result of 9/11.
<br><br>
To make sure that the increase in hate crimes in 2001 was actually due to 9/11, we first had to verify that other hate crime biases did not change significantly during that time. Graphing anti-Black and anti-White hate crimes over time verified that those kinds of biases did not change, and that the increase in 2001 was due to increased anti-Islam and anti-Arab hate crimes.

<center>
```{r out.width=c('50%', '50%'), fig.show='hold', echo=FALSE}
antiblack2001
antiwhite2001
```
</center>

Instead, we found there was a significant spike in the number of Anti-Islam and Anti-Arab hate crimes following 9/11.

<center>
```{r out.width=c('50%', '50%'), fig.show='hold', echo=FALSE}
antiarab2001
antiislam2001
```
</center>

<br><br>
Then, to further explore what the causes might have been, we found the counts of incidents of hate crime for each month, and discovered a few important trends.
<br><br>
First, we found that the frequencies of hate crime on a national scale is very cyclical, typically increasing during spring and decreasing sharply as we approach the end of the year, presumably because of winter or because of end-of-year festivities such as Thanksgiving and Christmas.

<center>
```{r ProjectAnalysis, echo=FALSE}

ggplot(data=data, aes(x=data$month, y=data$count, group=1)) + geom_line() + 
  scale_x_date(date_breaks = "years" , date_labels = "%Y") + labs(title="National Incidents of Hate Crime over Time", x="Date (year)", y = "National Hate Crimes Count") +
  theme(plot.title = element_text(hjust = 0.5)) + geom_vline(aes(xintercept=terr), color="blue", linetype="dashed", size=0.5) + geom_vline(aes(xintercept=trump), color="red", linetype="dashed", size=0.5)

```
</center>

Secondly, from the graph, we discovered a large spike in hate crimes near the end of the year 2001. The dashed blue line is the month of 9/11, which suggests that (in the U.S., not including the terrorist attack) the reason for the *significant* spike is due to 9/11. Additionally, we noticed that beginning from 2016, the incidents of hate crime no longer followed the generally decreasing but somewhat cyclical counts of hate crime. The dashed red line is the month that Trump was elected as president, and we also observe a spike that month. 
<br><br>
Additionally, we did not see it drop as much as we would have expected (given that the number of incidents drops towards the end of the year).

## Relationships between Anti-Black or African American and Anti-White Hate Crimes

We also plotted the number of anti-White and anti-Black hate crimes over time and did a linear regression to see if there was a general trend in the number of hate crimes. 

<center>
```{r out.width=c('50%', '50%'), fig.show='hold', echo=FALSE}

ggplot(data=data.frame(year, AfCrimes), aes(x=year,y=AfCrimes)) + geom_point(stat="identity") + geom_abline(aes(slope = af.b, intercept = af.a)) + labs(title = "Anti-Black or African American Hate Crimes Over Time", x = "Year", y = "Hate Crimes") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(data=data.frame(year, WhCrimes), aes(x=year,y=WhCrimes)) + geom_point(stat="identity") + geom_abline(aes(slope = wh.b, intercept = wh.a)) + labs(title = "Anti-White Hate Crimes Over Time", x = "Year", y = "Hate Crimes") +
  theme(plot.title = element_text(hjust = 0.5))

```
</center>

The slope for the anti-Black hate crime linear regression was -70.96 and the intercept was 144878.9. 

```{r}
# Anti-Black or African American Hate Crimes
# intercept, slope
af.a; af.b
# y = -70.96 x + 144878.88
```

The slope for the anti-White hate crime linear regression was -12.9 and the intercept was 26633.25. 

```{r}
# Anti-White Hate Crimes
# intercept, slope
wh.a; wh.b
# y = -70.96 x + 144878.88
```

While anti-White hate crimes are decreasing at a slower rate, there were a lot more anti-Black hate crimes to begin with. If this general trend continues, the number of anti-White and anti-Black hate crimes should be about equal in the year 2036. However, the number of hate crimes has been increasing the past few years, so it is unclear if the trends found in our linear regressions will hold for future years.

## Hate vs. Hate
One theory that we considered is that places with more occurences of anti-Black hate crime would be generally areas that are white-dominated, and therefore, those areas would have less occurences of anti-White hate crime. Similarly, areas with more anti-White hate crime would have less anti-Black hate crime. To test this hypothesis, we created a matrix, where one column was a vector of the number of anti-Black hate crimes, one was of the anti-White hate crimes, one was of the anti-Asian hate crimes, one was of anti-Hispanic or Latino hate crimes, and the last one was of anti-American Indian hate crimes. In our matrix, we also made sure to only include data from cities that had a population from 250,000 to 499,999, just to ensure that the number of hate crimes are not just due to the size of the overall population. Creating a correlation matrix from those vectors shows that the strongest correlation of the different hate crime biases was between anti-White and anti-Black hate crimes with a correlation of 0.6. Contrary to what we initially thought, anti-White and anti-Black hate crimes are positively correlated with each other. Perhaps in areas with high racial tensions, "hate" in general is larger, affecting all racial groups. 


## Variation of Hate Crimes Across the U.S.

We also wanted to see if the frequency of hate crime varied substantially across the U.S. We didn't want to simply plot the number of hate crimes that occured in an area because states with larger populations to begin with would be expected to have a higher number of hate crimes. Therefore, we merged our FBI hate crime data with population data of cities from the 2010 U.S. Census. For each state, we took the number of hate crimes from all the cities in that state divided by the sum of the population of those cities, giving us the rate of hate crimes per person. Mapping the results onto a map of the U.S. (with lighter colors indicating higher rates of hate crimes) shows us some patterns among the states. 

<br>
<center>
```{r, echo=FALSE}
if (require(maps)) {
  states_map <- map_data("state")
  ggplot (hatecrimes, aes(map_id = state)) + 
    geom_map (aes(fill = State_Rate), map = states_map) + 
    expand_limits(x = states_map$long, y= states_map$lat) + 
    labs(title="Rates of Hate Crimes per State (per 100)", x="", y= "") + theme(plot.title = element_text(hjust = 0.5))
}
```
</center>

The states that stand out in particular are South Dakota, New York, and Mississippi, which are shaded with the lightest colors (most hate crimes given the population). Initially, before starting this project, we had hypothesized that states in the South would have the worst rates of hate crime (especially given the historical tendencies). 

Unfortunately, the population data from the Census was not merged perfectly with the FBI crime data, and we wanted to make sure that we had as much available data on hate crime in cities as possible. Luckily, the hate crime dataset also classifies cities/counties by population. Therefore, by only comparing the total number of hate crimes for cities of a certain size (we tested it for cities with a population from 250,000 to 499,999) we would ensure that the overall populations of those cities are comparable. Surprisingly, for cities (with a population from 250,000 to 499,999), Southern cities actually had a lower mean for the total number of hate crimes than the rest of the U.S. 

From the map, the Midwest also seemed to have relatively higher rates of hate crimes, so we decided to compare Southern cities directly with Northeastern cities. Selecting a subset of the data that only contained cities for the Northeast and for the South, we ran a permutation test. The p-value was above 0.5, showing that surprisingly, the South did not actually have statistically higher hate crime rates than the Northeast. 

As we mentioned previously, the Midwestern states seemed to have relatively higher rates of hate crime, so we did another permutation test comparing Midwestern cities with cities in the other regions (Northeast, South, West) of the same population size. The p-value was about 0.008, showing that cities in the Midwest do have statistically higher hate crime rates. A chi-square test returns a similar p-value. 

```{r}
chi
```

<center>
``` {r, echo=FALSE}
x <- c(0,8)
c5 <- function (x) {dchisq(x, df = 1)}
ggplot(data.frame(x), aes(x=x)) + labs(title="Chi-Squared Distribution (df = 1)", x="chi-squared statistic", y = "p-value") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  stat_function(
    fun = c5
  ) + geom_vline(aes(xintercept=chi), color="red", linetype="dashed", size=0.5)
```
</center>

The chi square statistic value was about 5.6, and for one degrees of freedom, the chi-square test returned a p-value of about 0.1

We were also curious to see the breakdown of the different hate crime biases (anti-Black, anti-Catholic, etc.) across the U.S. One breakdown that was interesting was mapping the anti-American Indian or Alaska Native hate crime rates.

<center>
``` {r, echo=FALSE}
if (require(maps)) {
  states_map <- map_data("state")
  ggplot (hatecrimes, aes(map_id = state)) + 
    geom_map (aes(fill = staterate), map = states_map) + 
    expand_limits(x = states_map$long, y= states_map$lat) + 
    labs(title="Rate of Anti-American Indian or Alaska Native Hate Crimes (per 100)", x="", y= "") + theme(plot.title = element_text(hjust = 0.5))
}
```
</center>

While most states had generally low rates of anti-American Indian crime rates (presumably also because of the relatively low percentage of American Indian populations in the U.S.), Kentucky stood out as having much higher rates. To test whether these higher rates were actually significant, a permutation test was conducted. We controlled for population size in the cities (by only comparing cities with a population from 250,000 to 499,999), and the permutation test return a p-value that was about 10^(-4), demonstrating a statistically significant difference between Kentucky and the other states. 


## Non-MSA vs MSA Counties

We wanted to compare whether the actual composition or qualities of an area could impact the number of hate crimes as well. Therefore, we decided to compare non-MSA vs. MSA counties. MSA stands for Micropolitan/Metropolitan Statistical Area, which are areas that tend to have areas with very dense populations and a "high degree of social and economic integration" with the surrounding areas (Federal Register). We wanted the population of the counties and the geographical layout of these non-MSA and MSA counties to be comparable as well to ensure that any hate crime differences found between them was truly due to their economic qualities or population densities instead of other outside factors. Therefore, we looked at non-MSA and MSA counties with populations from 25,000 to 99,999. Calculating the fraction of non-MSA and MSA counties in the Northeast, South, West, and Midwest shows that the distribution of non-MSA and MSA counties in the different regions are similar as well. A permutation test was conducted for the total number of hate crimes for MSA and non-MSA counties. The p-value was about 0.049, which demonstrates that MSA counties have statistically higher levels of hate crime than non-MSA counties. A chi-square test confirms that the result is indeed statistically significant. Although the non-MSA and MSA counties compared in our analysis have similar populations, the fact that MSA counties have areas with higher population densities perhaps allow for more interactions between people and for more hate crimes to happen.

```{r}
msachi
```

<center>
``` {r, echo=FALSE}
x <- c(0,8)
c5 <- function (x) {dchisq(x, df = 1)}
ggplot(data.frame(x), aes(x=x)) + labs(title="Chi-Squared Distribution (df = 1)", x="chi-squared statistic", y = "p-value") +
  theme(plot.title = element_text(hjust = 0.5)) + 
  stat_function(
    fun = c5
  ) + geom_vline(aes(xintercept=msachi), color="red", linetype="dashed", size=0.5)
```
</center>


### Sources
https://www.fbi.gov/services/cjis/ucr/hate-crime

https://www.census.gov/data/tables/2017/demo/popest/total-cities-and-towns.html

https://www.govinfo.gov/content/pkg/FR-2010-06-28/pdf/2010-15605.pdf

`Analysis performed by Jenny Gu and Yijiang Zhao`
